#!/usr/bin/env Rscript

args <- commandArgs()
fname <- normalizePath(sub("--file=", "", args[grepl("--file=", args)]))
thisfile <- readLines(fname)
newfname <- paste0(tempdir(), "/", basename(fname))
writeLines(thisfile[-1:-which(thisfile == "q(\"no\")")], newfname)


library("optparse")
option_list = list(
	make_option(c("--rds"), type="character", default="", 
		help="rds file name", metavar="character")
); 
opt_parser = OptionParser(option_list=option_list);
opt = parse_args(opt_parser, positional_arguments=TRUE);
print(opt)
#print(opt$args)

#opt$args[1] = normalizePath(opt$args[1])
#print(opt$args)
#print(opt$args[1])

#print(opt$options$rds)
#opt$options$rds = normalizePath(opt$options$rds)
#print(opt$options$rds)
#output_base = paste( gsub(".RDS","",opt$options$rds) , "median", sep=".")

print(opt$args[1])
opt$args[1] = normalizePath(opt$args[1])
print(opt$args[1])
output_base = paste( gsub(".RDS","",opt$args[1]) , "median", sep=".")
output_file = paste(output_base, "html", sep=".")
print(output_file)

rmarkdown::render(newfname, output_dir = dirname(output_file), output_file = output_file )

q("no")




---
title: "REdiscoverTE Median Plotter"
author: "Jake W"
date: "2024-05-01"
output: html_document
fig_width: 12
fig_height: 4
---


```{r args}
args
```


```{r "Figure Settings", include = FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=6) 
```



```{r}
#r=readRDS( opt$options$rds )
r=readRDS( opt$args[1] )
r=r$counts
```

```{r}
dim(r)
```


```{r}
medians=apply(r,1,median)
head(medians)
```


```{r}
min(medians)
```

```{r}
median(medians)
```

```{r}
max(medians)
```


```{r}
hist(medians,
breaks=25,
main="Median Histogram",
xlab="Median"
)
```



About a half a dozen are above 10,000.

2 are above 100,000 really screwing with the spread.

```{r}
tail(sort(medians),24)
```



```{r}
hist(medians,
breaks=25,
ylim=c(0,10),
main="Median Histogram",
xlab="Median"
)
```

Basically, all except 2 are in one bar. Not helpful.


```{r}

length(medians[which(medians==0)])

```


Drop all of the simple repeats ( in the format of "(....)n" ).

Many are 0.

```{r}
r=r[!grepl(")n$", row.names(r)),]
dim(r)
```

Many are still 0.

Drop those 2 high values.

```{r}
r=r[!grepl("SSU-rRNA_Hsa", row.names(r)),]
r=r[!grepl("LSU-rRNA_Hsa", row.names(r)),]
```

```{r}
medians=apply(r,1,median)
```

```{r}

length(medians[which(medians==0)])

```


```{r}
min(medians)
```

```{r}
median(medians)
```

```{r}
max(medians)
```



```{r}
hist(medians,
breaks=100,
main="Median Histogram",
xlab="Median"
)
```


```{r}
hist(medians,
breaks=100,
ylim=c(0,10),
main="Median Histogram",
xlab="Median"
)
```


Reset

```{r}
r=readRDS( opt$args[1] )
r=r$counts
r=r[!grepl(")n$", row.names(r)),]
medians=apply(r,1,median)
medians=medians[which(medians>0)]
```



```{r}

length(medians)

```


```{r}

length(medians[which(medians==0)])

```


```{r}
hist(medians[which(medians <3000)],
breaks=100,
ylim=c(0,50),
main="Median Histogram",
xlab="Median"
)
```


```{r}
hist(medians[which(medians <3000)],
breaks=100,
ylim=c(0,100),
main="Median Histogram",
xlab="Median"
)
```


```{r}
hist(medians[which(medians <3000)],
breaks=100,
ylim=c(0,500),
main="Median Histogram",
xlab="Median"
)
```


```{r}
hist(medians[which(medians <3000)],
breaks=100,
ylim=c(0,1000),
main="Median Histogram",
xlab="Median"
)
```

```{r}
hist(medians[which(medians <300)],
breaks=100,
 xlim=c(0,300),
main="Median Histogram",
xlab="Median"
)
```

